R for Clinical Study Reports and Submission

### 2022-12-07 submission-package.Rmd

```{r, echo=FALSE}
library("kableExtra")

df1 <- data.frame(
  program = c(
    "tlf-01-disposition.txt",
    "tlf-02-population.txt",
    "tlf-03-baseline.txt",
    "tlf-04-efficacy.txt",
    "tlf-05-ae-summary.txt",
    "tlf-06-ae-spec.txt"
  ),
  table = c(
    "Table x.y.z"
  ),
  desc = c(
    "Disposition of Patients",
    "Participants Accounting in Analysis Population (All Participants Randomized)",
    "Participant Baseline Characteristics (All Participants Randomized)",
    "ANCOVA of Change from Baseline Glucose (mmol/L) at Week 24 <br> LOCF <br> Efficacy Analysis Population",
    "Analysis of Adverse Event Summary (Safety Analysis Population)",
    "Analysis of Participants With Specific Adverse Events (Safety Analysis Population)"
  ),
  stringsAsFactors = FALSE
)

names(df1) <- c("Program Name", "Output Table", "Title")

df1 %>%
  kbl(escape = FALSE) %>%
  kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
  column_spec(1, extra_css = "border: 1px solid #000; text-align: center;", width = "18rem") %>%
  column_spec(2, extra_css = "border: 1px solid #000; text-align: center;", width = "16rem") %>%
  column_spec(3, extra_css = "border: 1px solid #000;", width = "28rem") %>%
  row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;")

df2 <- data.frame(
  package = c("pkglite", "haven", "dplyr", "tidyr", "emmeans", "r2rtf"),
  version = NA,
  description = c(
    "Prepare submission package",
    "Read SAS datasets",
    "Manipulate datasets",
    "Manipulate datasets",
    "Least-squares means estimation",
    "Create RTF tables"
  ),
  stringsAsFactors = FALSE
)

@@

names(df2) <- c(
  "Open-Source R Analysis Package",
  "Package Version",
  "Analysis Package Description"
)

df2 %>%
  kbl() %>%
  kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
  column_spec(1, extra_css = "border: 1px solid #000; text-align: center;", width = "18rem") %>%
  column_spec(2, extra_css = "border: 1px solid #000; text-align: center;", width = "16rem") %>%
  column_spec(3, extra_css = "border: 1px solid #000;", width = "28rem") %>%
  row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;") %>%
  footnote(general = "R package versions based on CRAN on 2021-08-06", general_title = "")

df3 <- data.frame(
  package = "esubdemo",
  version = "0.1.0",
  description = "A demo package for analysis and reporting of clinical trials",
  stringsAsFactors = FALSE
)

names(df3) <- c(
  "Proprietary R Analysis Package",
  "Package Version",
  "Analysis Package Description"
)

df3 %>%
  kbl() %>%
  kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
  column_spec(1, extra_css = "border: 1px solid #000; text-align: center;", width = "18rem") %>%
  column_spec(2, extra_css = "border: 1px solid #000; text-align: center;", width = "16rem") %>%
  column_spec(3, extra_css = "border: 1px solid #000;", width = "28rem") %>%
  row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;")
```

@@

names(df4) <- c(
  "Table Reference",
  "Table Title",
  "Programming Language",
  "Program Name (programs)",
  "Input File Name / Analysis (datasets)"
)

df4 %>%
  kbl() %>%
  kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
  column_spec(1, extra_css = "border: 1px solid #000; text-align: center;") %>%
  column_spec(2, extra_css = "border: 1px solid #000; text-align: center;") %>%
  column_spec(3, extra_css = "border: 1px solid #000; text-align: center;") %>%
  column_spec(4, extra_css = "border: 1px solid #000; text-align: center;") %>%
  column_spec(5, extra_css = "border: 1px solid #000; text-align: center;") %>%
  row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;")
```

@@

```{r, echo=FALSE}
df5 <- data.frame(
  "x1" = c(
    "Table Reference: [Ref. x.y.z: P001ZZZ9999: Table 1-1]",
    "Analysis Result",
    "Analysis Parameters (s)",
    "Analysis Reason",
    "Analysis Purpose",
    "...",
    "Programming Statements"
  ),
  "x2" = c(
    "...",
    "...",
    "...",
    "...",
    "...",
    "...",
    "(R version 4.1.1), [P001ZZZ9999: programs-tlf-01-disposition]"
  ),
  stringsAsFactors = FALSE
)

df5 %>%
  kbl(format = "html") %>%
  kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
  column_spec(1, extra_css = "border: 1px solid #000;", bold = TRUE) %>%
  column_spec(2, extra_css = "border: 1px solid #000;") %>%
  row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000;") %>%
  gsub("<thead>.*</thead>", "", .)
```

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library("kableExtra")
```

```{r, echo=FALSE}
df1 <- data.frame(
  program = c(
    "tlf-01-disposition.txt",
    "tlf-02-population.txt",
    "tlf-03-baseline.txt",
    "tlf-04-efficacy.txt",
    "tlf-05-ae-summary.txt",
    "tlf-06-ae-spec.txt"
  ),
  table = c(
    "Table x.y.z"
  ),
  desc = c(
    "Disposition of Patients",
    "Participants Accounting in Analysis Population (All Participants Randomized)",
    "Participant Baseline Characteristics (All Participants Randomized)",
    "ANCOVA of Change from Baseline Glucose (mmol/L) at Week 24 <br> LOCF <br> Efficacy Analysis Population",
    "Analysis of Adverse Event Summary (Safety Analysis Population)",
    "Analysis of Participants With Specific Adverse Events (Safety Analysis Population)"
  ),
  stringsAsFactors = FALSE
)

names(df1) <- c("Program Name", "Output Table", "Title")
```

```{r, echo=FALSE}
if (knitr::is_latex_output()) {
  df1 %>% kbl(format = "latex")
}

if (knitr::is_html_output()) {
  df1 %>%
    kbl(format = "html", escape = FALSE) %>%
    kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
    column_spec(1, extra_css = "border: 1px solid #000; text-align: center;", width = "18rem") %>%
    column_spec(2, extra_css = "border: 1px solid #000; text-align: center;", width = "16rem") %>%
    column_spec(3, extra_css = "border: 1px solid #000;", width = "28rem") %>%
    row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;")
}
```

```{r, echo=FALSE}
df2 <- data.frame(
  package = c("pkglite", "haven", "dplyr", "tidyr", "emmeans", "r2rtf"),
  version = NA,
  description = c(
    "Prepare submission package",
    "Read SAS datasets",
    "Manipulate datasets",
    "Manipulate datasets",
    "Least-squares means estimation",
    "Create RTF tables"
  ),
  stringsAsFactors = FALSE
)

@@

names(df2) <- c(
  "Open-Source R Analysis Package",
  "Package Version",
  "Analysis Package Description"
)
```

```{r, echo=FALSE}
if (knitr::is_latex_output()) {
  df2 %>% kbl(format = "latex")
}

if (knitr::is_html_output()) {
  df2 %>%
    kbl(format = "html") %>%
    kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
    column_spec(1, extra_css = "border: 1px solid #000; text-align: center;", width = "18rem") %>%
    column_spec(2, extra_css = "border: 1px solid #000; text-align: center;", width = "16rem") %>%
    column_spec(3, extra_css = "border: 1px solid #000;", width = "28rem") %>%
    row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;") %>%
    footnote(general = "R package versions based on CRAN on 2021-08-06", general_title = "")
}
```

```{r, echo=FALSE}
df3 <- data.frame(
  package = "esubdemo",
  version = "0.1.0",
  description = "A demo package for analysis and reporting of clinical trials",
  stringsAsFactors = FALSE
)

names(df3) <- c(
  "Proprietary R Analysis Package",
  "Package Version",
  "Analysis Package Description"
)
```

```{r, echo=FALSE}
if (knitr::is_latex_output()) {
  df3 %>% kbl(format = "latex")
}

if (knitr::is_html_output()) {
  df3 %>%
    kbl(format = "html") %>%
    kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
    column_spec(1, extra_css = "border: 1px solid #000; text-align: center;", width = "18rem") %>%
    column_spec(2, extra_css = "border: 1px solid #000; text-align: center;", width = "16rem") %>%
    column_spec(3, extra_css = "border: 1px solid #000;", width = "28rem") %>%
    row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;")
}
```

@@

names(df4) <- c(
  "Table Reference",
  "Table Title",
  "Programming Language",
  "Program Name (programs)",
  "Input File Name / Analysis (datasets)"
)
```

```{r}
if (knitr::is_latex_output()) {
  df4 %>% kbl(format = "latex")
}

if (knitr::is_html_output()) {
  df4 %>%
    kbl(format = "html") %>%
    kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
    column_spec(1, extra_css = "border: 1px solid #000; text-align: center;") %>%
    column_spec(2, extra_css = "border: 1px solid #000; text-align: center;") %>%
    column_spec(3, extra_css = "border: 1px solid #000; text-align: center;") %>%
    column_spec(4, extra_css = "border: 1px solid #000; text-align: center;") %>%
    column_spec(5, extra_css = "border: 1px solid #000; text-align: center;") %>%
    row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000; text-align: center;")
}
```

@@

```{r, echo=FALSE}
df5 <- data.frame(
  "x1" = c(
    "Table Reference: [Ref. x.y.z: P001ZZZ9999: Table 1-1]",
    "Analysis Result",
    "Analysis Parameters (s)",
    "Analysis Reason",
    "Analysis Purpose",
    "...",
    "Programming Statements"
  ),
  "x2" = c(
    "...",
    "...",
    "...",
    "...",
    "...",
    "...",
    "(R version 4.1.1), [P001ZZZ9999: programs-tlf-01-disposition]"
  ),
  stringsAsFactors = FALSE
)
```

```{r, echo=FALSE}
if (knitr::is_latex_output()) {
  df5 %>% kbl(format = "latex")
}

if (knitr::is_html_output()) {
  df5 %>%
    kbl(format = "html") %>%
    kable_classic(full_width = FALSE, html_font = "'Times New Roman', Times, serif", font_size = 18) %>%
    column_spec(1, extra_css = "border: 1px solid #000;", bold = TRUE) %>%
    column_spec(2, extra_css = "border: 1px solid #000;") %>%
    row_spec(0, background = "#DFDFDF", bold = TRUE, extra_css = "border: 1px solid #000;") %>%
    gsub("<thead>.*</thead>", "", .)
}
```

### 2022-12-07 tlf-ae-summary.Rmd

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_ae_summary.pdf")
```

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_ae_summary.pdf")
```

>>>>

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_ae_summary.pdf")
```

@@

```{r, include=FALSE}
rtf2pdf("tlf/tlf_ae_summary.rtf")
```

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_ae_summary.pdf")
```

### 2022-12-07 tlf-assemble.Rmd
Update tlf-assemble.Rmd

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/rtf-combine.pdf")
```

@@

::: {.rmdnote}
There is a slight learning curve to learning how toggle fields work 
in Microsoft Words. 
After you become familiar with the workflow, 
toggle fields can extend your capability to manage a large amount of TLFs
in RTF format.
:::

>>>>

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/rtf-combine.pdf")
```

@@

::: {.rmdnote}
There is a slight learning curve on how toggle fields work 
in Microsoft Words. 
After you become familiar with the workflow, 
toggle fields can extend your capability to manage a large amount of TLFs
in RTF format.
:::

### 2022-12-07 tlf-baseline.Rmd
minor updates

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_base.pdf")
```

@@

`text_indent_first` and `text_indent_left` are used to control the indent
space of text. They are helpful when you need to control the white space
of a long phrase "AMERICAN INDIAN OR ALASKA NATIVE"
in the table provides an example.

>>>>

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_base.pdf")
```

@@

`text_indent_first` and `text_indent_left` are used to control the indent
space of text. They are helpful when you need to control the white space
of a long phrase, "AMERICAN INDIAN OR ALASKA NATIVE"
in the table provides an example.

### 2022-12-07 tlf-disposition.Rmd

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_disp.pdf")
```

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_disp.pdf")
```

>>>>

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_disp.pdf")
```

@@

```{r, include=FALSE}
rtf2pdf("tlf/tbl_disp.rtf")
```

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_disp.pdf")
```

### 2022-12-07 tlf-efficacy-ancova.Rmd
rename file

# Efficacy

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff.pdf")
```

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff1.pdf")
```

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff2.pdf")
```

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff.pdf")
```

>>>>

# Efficacy table

@@

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff.pdf")
```

@@

```{r, include=FALSE}
rtf2pdf("tlf/tlf_eff1.rtf")
```

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff1.pdf")
```

@@

```{r, include=FALSE}
rtf2pdf("tlf/tlf_eff2.rtf")
```

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff2.pdf")
```

@@

```{r, include=FALSE}
rtf2pdf("tlf/tlf_eff.rtf")
```

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_eff.pdf")
```

### 2022-12-07 tlf-population.Rmd
source common.R in each Rmd to allow compile individually
fix some minor issues

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

@@

The function can be enhanced in multiple ways, but here we only focus on
simplicity and readability. More details about writing R functions can
be found in the [Stat454
course](https://stat545.com/functions-part1.html).

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

>>>>

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

@@

The function can be enhanced in multiple ways, but here we only focus on
simplicity and readability. More details about writing R functions can
be found in the [STAT 545
course](https://stat545.com/functions-part1.html).

@@

```{r, include=FALSE}
rtf2pdf("tlf/tbl_pop.rtf")
```

```{r, out.width = "100%", out.height = if (knitr::is_html_output()) "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```
